Monte Carlo tree search compared to A\(^*\) in airspace configuration decision problem

### 2018-05-20 article/article.tex
(+) Introduced new results

\subsection{Accuracy}
Accuracy is assessed by comparing the outcome of the MCTS algorithm with the
A\(^*\) cost and a greedy algorithm cost.
The time allowed to the MCTS to run is considered long enough to allow the MCTS
to converge to its best solution. The results are summarised in
table~\ref{tab:methods_compare}.

The graph~\ref{fig:grouping} asserts the correct behaviour of both algorithms
A\(^*\) and MCTS regarding grouping and degrouping. On the \(y\) axis is
indicated the number of sectors. The trafic activity is the following,
\begin{enumerate}
  \item\label{scen:busy}
    The first part of the scenario shows a busy situation, all sectors are
    overloaded;
  \item\label{scen:quiet}
    then follows a quiet moment with few aircraft, much of the sectors are
    underloaded;
  \item\label{scen:partial}
    the end of the scenario has two overload modules and the rest under
    normal or under load.
\end{enumerate}
The expected behaviour is a tendency to increase the number of sectors through
splits to cope with part~\ref{scen:busy} followed by many merges to reduce the
number of sectors in response to phase~\ref{scen:quiet}. The algorithms should
end with a specific sector created, the one overloaded in~\ref{scen:partial} and
the rest grouped. Seeing~\ref{fig:grouping}, both algorithm behave as expected.
\begin{table}
  \centering
  \begin{tabulary}
    {\textwidth}{LRR}
    Alg. & Cost & Err.\\\toprule
    A\(^*\) & 77.31 & 0\%    \\
    MCTS    & 82.48 & 6.7\%  \\
    Greedy  & 90.50 & 17.1\% \\
  \end{tabulary}
  \caption{Comparison of several methods on the generated
  scenario.}\label{tab:methods_compare}
\end{table}

\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=timestep, ylabel=num of sectors,
      legend entries={A\(^*\), MCTS}]
      \foreach \i in {0,1}{%
        \addplot+ table [
          x expr={\lineno}, y index=\i
          ]
          {./data/grouping.data};
      }
    \end{axis}
  \end{tikzpicture}
  \caption{Grouping profile}\label{fig:grouping}
\end{figure}
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=time(s), ylabel=cost]
      \addplot+ table {data/shortterm.data};
    \end{axis}
  \end{tikzpicture}
  \caption{Monte Carlo tree search called with varying time allowed to compute
  each step, with a depth of 40 and \(\theta = 11\). The cost seems to be
  approximately stable from \(t=0.5\)s.}\label{fig:shortterm}
\end{figure}

@@

The converging speed may also depend on the depth of the tree searched through.
The figure~\ref{fig:timedepth} shows

\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=depth, ylabel=cost]
      \foreach \i in {1, 2} {%
        \addplot+ table [ x index = 0, y index = \i ] {./data/cmpdepth.data};
      }
    \end{axis}
  \end{tikzpicture}
  \caption{Cost difference for varying depth with 0.5 seconds allowed to MCTS,
  \(\theta=3\)}\caption{fig:cmpdepth}
\end{figure}

>>>>

\subsection{Accuracy}
The graph~\ref{fig:grouping} asserts the correct behaviour of both algorithms
A\(^*\) and MCTS regarding grouping and degrouping. On the \(y\) axis is
indicated the number of sectors. The trafic activity is the following,
\begin{enumerate}
  \item\label{scen:busy}
    The first part of the scenario shows a busy situation, all sectors are
    overloaded;
  \item\label{scen:quiet}
    then follows a quiet moment with few aircraft, much of the sectors are
    underloaded;
  \item\label{scen:partial}
    the end of the scenario has two overload modules and the rest under
    normal or under load.
\end{enumerate}
The expected behaviour is a tendency to increase the number of sectors through
splits to cope with part~\ref{scen:busy} followed by many merges to reduce the
number of sectors in response to phase~\ref{scen:quiet}. The algorithms should
end with a specific sector created, the one overloaded in~\ref{scen:partial} and
the rest grouped. Seeing~\ref{fig:grouping}, both algorithm behave as expected.
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=timestep, ylabel=num of sectors,
      legend entries={A\(^*\), MCTS}]
      \foreach \i in {0,1}{%
        \addplot+ table [
          x expr={\lineno}, y index=\i
          ]
          {./data/grouping.data};
      }
    \end{axis}
  \end{tikzpicture}
  \caption{Grouping profile}\label{fig:grouping}
\end{figure}

Accuracy can be quantitatively assessed by comparing the outcome of the MCTS
algorithm against the A\(^*\) cost and a greedy algorithm cost.
The time allowed to the MCTS to run is considered long enough to allow the MCTS
to converge to its best solution. The results are summarised in
table~\ref{tab:methods_compare}.
\begin{table}
  \centering
  \begin{tabulary}
    {\textwidth}{LRR}
    Alg.    & Cost  & Err.   \\
    \toprule
    A\(^*\) & 77.31 & 0\%    \\
    MCTS    & 82.48 & 6.7\%  \\
    Greedy  & 90.50 & 17.1\% \\
  \end{tabulary}
  \caption{Comparison of several methods on the generated
  scenario.}\label{tab:methods_compare}
\end{table}

The graph~\ref{fig:cmpdepth} shows the evolution of the difference of costs with
varying depth. The error rate is expected to grow as the depth increase since
the MCTS might fail to explore the best nodes among the exponentially growing
amount of them.
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=depth, ylabel=cost,
      legend entries = {A\(^*\), MCTS}]
      \foreach \i in {0, 1} {%
        \addplot+ table [
          x expr = {\lineno + 1}, y index = \i ] {./data/cmpdepth.data};
      }
    \end{axis}
  \end{tikzpicture}
  \caption{Cost difference for varying depth with 0.5 seconds allowed to
    MCTS.}\label{fig:cmpdepth}
\end{figure}

@@

The converging speed may also depend on the depth of the tree searched through.
The figure~\ref{fig:timedepth} shows
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=time(s), ylabel=cost]
      \addplot+ table {data/shortterm.data};
    \end{axis}
  \end{tikzpicture}
  \caption{Monte Carlo tree search called with varying time allowed to compute
  each step, with a depth of 40 and \(\theta = 11\). The cost seems to be
  approximately stable from \(t=0.5\)s.}\label{fig:shortterm}
\end{figure}
